Supplemental Material: Multidimensional analysis and detection of informative features in human brain white matter

### 2021-04-20 reports/manuscript/paper.tex
A few small things.

The $\beta$ coefficients exhibit high bundle level sparsity; only some
bundles are important, which can be be confirmed by observing the value of
$\alpha$, the regularization hyperparameter that controls the mixture of the
Group Lasso and lasso penalties, selected through nested cross-validation
(see \cref{eq:sgl}).
If $\alpha$ is closer to zero, it indicates that the phenotype in question
preferentially correlates with only a few groups of covariates. For the ALS
dataset, the SGL model has $\alpha = $ \alsLRatio
and the PCR-SGL model has \protect$\alpha = $ \protect\alsLRatioGpca,
confirming that the white matter correlates of ALS reside mostly in one
bundle, namely the CST.

@@

To test the performance of SGL in a continuous
regression task, we focus here on the prediction of biological age in three
datasets named WH, HBN, and Cam-CAN (see the methods section for a description of each dataset). Prediction of ``brain
age'' is a commonly undertaken task in neuroimaging machine learning, in part
because these predictions, and deviations therefrom, may be diagnostic of
overall brain health (for a review, see Cole et al.~\cite{Cole2019-rz}).
However, as Nelson et al.~\cite{nelson2019biomarkers} have observed, aging
biomarkers are subject to unique challenges and tend to be noncausative. Our
interest in aging here relies on its utility as a methodological benchmark.
Biological age operates on a natural scale, with meaningful and easily
understood units, and it's popularity as a machine learning target makes it
valuable for comparisons with other studies.

@@

Another limitation of our approach is in its computational execution time,
which, in our experience, is roughly five times slower than simpler models
like the lasso. This slow-down is likely to be insignificant for ``production''
cycles in which researchers report their findings, but may be prohibitive
for ``development'' cycles in which researchers repeatedly train models while
adjusting their analysis pipeline. In practice, we circumvent this limitation
by developing our analysis infrastructure with fast, simple models, and substituting
slower, more performant models afterward.

>>>>

The $\beta$ coefficients exhibit high bundle level sparsity; only some
bundles are important, which can be be confirmed by observing the value of
$\alpha$, the regularization hyperparameter that controls the mixture of the
Group Lasso and lasso penalties, selected through nested cross-validation
(see \cref{eq:sgl}).
If $\alpha$ is closer to zero, it indicates that the phenotype in question
preferentially correlates with only a few groups of covariates. For the ALS
dataset, the SGL model has $\alpha = $ \alsLRatio
~and the PCR-SGL model has \protect$\alpha = $ \protect\alsLRatioGpca,
confirming that the white matter correlates of ALS reside mostly in one
bundle, namely the CST.

@@

To test the performance of SGL in a continuous
regression task, we focus here on the prediction of biological age in three
datasets named WH, HBN, and Cam-CAN (see the Methods section for a description of each dataset). Prediction of ``brain
age'' is a commonly undertaken task in neuroimaging machine learning, in part
because these predictions, and deviations therefrom, may be diagnostic of
overall brain health (for a review, see Cole et al.~\cite{Cole2019-rz}).
However, as Nelson et al.~\cite{nelson2019biomarkers} have observed, aging
biomarkers are subject to unique challenges and tend to be noncausative. Our
interest in aging here relies on its utility as a methodological benchmark.
Biological age operates on a natural scale, with meaningful and easily
understood units, and it's popularity as a machine learning target makes it
valuable for comparisons with other studies.

@@

Another limitation of our approach is in its computational execution time,
which, in our experience, is roughly five times slower than simpler models
like the lasso. This slow-down is likely to be insignificant for ``production''
cycles in which researchers report their findings, but may be prohibitive
for ``development'' cycles in which researchers repeatedly train models while
adjusting their analysis pipeline. In practice, we circumvent this limitation
by developing our analysis infrastructure with fast, simple models, and substituting
slower, more performant models afterwards. The consistent \texttt{Scikit Learn}-based application programming interface of \texttt{AFQ-Insight} provides a good basis for this pattern.

### 2021-04-20 reports/manuscript/paper_changes.tex
A few small things.

The $\beta$ coefficients exhibit high bundle level sparsity; only some
bundles are important, which can be be confirmed by observing the value of
$\alpha$, the regularization hyperparameter that controls the mixture of the
Group Lasso and lasso penalties, selected through nested cross-validation
(see \cref{eq:sgl}).
If $\alpha$ is closer to zero, it indicates that the phenotype in question
preferentially correlates with only a few groups of covariates. For the ALS
dataset, \added{the SGL model has} $\alpha = $ \alsLRatio
\added{and the PCR-SGL model has \protect$\alpha = $ \protect\alsLRatioGpca},
confirming that the white matter correlates of ALS reside mostly in one
bundle, namely the CST.

@@

To test the performance of SGL \deleted{with tractometry data} in a continuous
regression task, we focus here on the prediction of biological age in three
datasets named WH, HBN, and Cam-CAN (see the methods section for a description of each dataset). Prediction of ``brain
age'' is a commonly undertaken task in neuroimaging machine learning, in part
because these predictions, and deviations therefrom, may be diagnostic of
overall brain health (for a review, see Cole et al.~\cite{Cole2019-rz}).
However, as Nelson et al.~\cite{nelson2019biomarkers} have observed, aging
biomarkers are subject to unique challenges and tend to be noncausative. Our
interest in aging here relies on its utility as a methodological benchmark.
Biological age operates on a natural scale, with meaningful and easily
understood units, and it's popularity as a machine learning target makes it
valuable for comparisons with other studies.

@@

To evaluate the fit of the model, we used a nested cross-validation
procedure. In this procedure, batches of subjects are held out. For each
batch (or fold), the model is fully fit without this data. Then, once the
parameters are fixed, the model is applied to predict the ages of held out
subjects based on the linear coeffiecients. This scheme automatically finds
the right level of sparsity and fits the coefficients to the ill-posed linear
model, while guarding against overfitting. SGL accurately predicts the age of
the subjects in this procedure, with a median absolute error of {\whMae},
{\hbnMae}, and {\camcanMae} years for the WH, HBN, and Cam-CAN datasets,
respectively and coefficients of determination $R^2 = $ {\whRsq} , {\hbnRsq},
and {\camcanRsq}, respectively (see \cref{fig:age-results}, top panels). The
predictions for Cam-CAN are competitive with a recent state of the art
prediction \cite{mcpherson2020single}, which used streamline density to
estimate the brain's structural connectivity and achieved $R^2 = 0.63$. The
median absolute errors are also lower than the results of a recent study that
predicted age in a large sample that included the Cam-CAN data, and was based
on diffusion MRI features \cite{Richard2018-ux}.

@@

\added{Another limitation of our approach is in its computational execution time,}
\added{which, in our experience, is roughly five times slower than simpler models}
\added{like the lasso. This slow-down is likely to be insignificant for ``production''}
\added{cycles in which researchers report their findings, but may be prohibitive}
\added{for ``development'' cycles in which researchers repeatedly train models while}
\added{adjusting their analysis pipeline. In practice, we circumvent this limitation}
\added{by developing our analysis infrastructure with fast, simple models, and substituting}
\added{slower, more performant models afterward.}

>>>>

The $\beta$ coefficients exhibit high bundle level sparsity; only some
bundles are important, which can be be confirmed by observing the value of
$\alpha$, the regularization hyperparameter that controls the mixture of the
Group Lasso and lasso penalties, selected through nested cross-validation
(see \cref{eq:sgl}).
If $\alpha$ is closer to zero, it indicates that the phenotype in question
preferentially correlates with only a few groups of covariates. For the ALS
dataset, \added{the SGL model has} $\alpha = $ \alsLRatio
~\added{and the PCR-SGL model has \protect$\alpha = $ \protect\alsLRatioGpca},
confirming that the white matter correlates of ALS reside mostly in one
bundle, namely the CST.

@@

To test the performance of SGL \deleted{with tractometry data} in a continuous
regression task, we focus here on the prediction of biological age in three
datasets named WH, HBN, and Cam-CAN (see the Methods section for a description of each dataset). Prediction of ``brain
age'' is a commonly undertaken task in neuroimaging machine learning, in part
because these predictions, and deviations therefrom, may be diagnostic of
overall brain health (for a review, see Cole et al.~\cite{Cole2019-rz}).
However, as Nelson et al.~\cite{nelson2019biomarkers} have observed, aging
biomarkers are subject to unique challenges and tend to be noncausative. Our
interest in aging here relies on its utility as a methodological benchmark.
Biological age operates on a natural scale, with meaningful and easily
understood units, and it's popularity as a machine learning target makes it
valuable for comparisons with other studies.

@@

To evaluate the fit of the model, we used a nested cross-validation
procedure. In this procedure, batches of subjects are held out. For each
batch (or fold), the model is fully fit without this data. Then, once the
parameters are fixed, the model is applied to predict the ages of held out
subjects based on the linear coeffiecients. This scheme automatically finds
the right level of sparsity and fits the coefficients to the ill-posed linear
model, while guarding against overfitting. SGL accurately predicts the age of
the subjects in this procedure, with a median absolute error of {\whMae},
\added{{\hbnMae}}, and {\camcanMae} years for the WH, HBN, and Cam-CAN datasets,
respectively and coefficients of determination $R^2 = $ {\whRsq} , \added{{\hbnRsq}}, and {\camcanRsq},
respectively (see \cref{fig:age-results}, top panels). The
predictions for Cam-CAN are competitive with a recent state of the art
prediction \cite{mcpherson2020single}, which used streamline density to
estimate the brain's structural connectivity and achieved $R^2 = 0.63$. The
median absolute errors are also lower than the results of a recent study that
predicted age in a large sample that included the Cam-CAN data, and was based
on diffusion MRI features \cite{Richard2018-ux}.

@@

\added{Another limitation of our approach is in its computational execution time,}
\added{which, in our experience, is roughly five times slower than simpler models}
\added{like the lasso. This slow-down is likely to be insignificant for ``production''}
\added{cycles in which researchers report their findings, but may be prohibitive}
\added{for ``development'' cycles in which researchers repeatedly train models while}
\added{adjusting their analysis pipeline. In practice, we circumvent this limitation}
\added{by developing our analysis infrastructure with fast, simple models, and substituting}
\added{slower, more performant models afterwards. The consistent \texttt{Scikit Learn}-based application programming interface of \texttt{AFQ-Insight} provides a good basis for this pattern.}
